---
title: "Reading Data from the web"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(httr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Scrape a table

I want the first table from [this page](http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm)

Read in the html:

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)
```

Extract the table(s); focus on the first one

```{r}
table_marij=
  drug_use_html %>% 
  html_nodes(css = "table") %>% 
  first() %>%
  html_table() %>% 
  slice(-1) %>% 
  as_tibble()
```

## Stars Wars movie info

I want the data from [here](https://www.imdb.com/list/ls070150896/)

```{r}
url = "https://www.imdb.com/list/ls070150896/"

swm_html = read_html(url)
```

Grab elements that I want:

```{r}
title_vec =
  swm_html %>%
  html_nodes(css = ".lister-item-header a") %>% 
  html_text()

gross_revenue_vec =
  swm_html %>%
  html_nodes(css = ".text-muted .ghost~ .text-muted+ span") %>% 
  html_text()

run_time_vec =
  swm_html %>%
  html_nodes(css = ".runtime") %>% 
  html_text()

swm_df =
  tibble(
    title = title_vec,
    gross_rev = gross_revenue_vec,
    runtime = run_time_vec)
```

## Get some water data

This is coming from an API:

```{r}
nyc_water =
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") %>% 
  content("parsed")

nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.json") %>% 
  content("text") %>%
  jsonlite::fromJSON() %>%
  as_tibble()
```

## BRFSS Dataset (Behavioral Risk Factors)

Same process, different data:

```{r}
brfss_2010 =
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv",
      query = list("$limit" = 5000)) %>%
  content("parsed")
```

## Some data aren't so nice

Looking at Pokemon:

```{r}
pokemon_data =
  GET("https://pokeapi.co/api/v2/pokemon/1") %>% 
  content

pokemon_data$name
pokemon_data$height
pokemon_data$abilities
```

## Closing thoughts

It's easy to make unreasonable requests in scraping and API use.
Remember that every time you knit this, you are taking things from websites. If you do this a lot, you could crash a server.